\chapter{Introduction}
\label{chap:intro}
\chaptermark{Optional running chapter heading}

Modern software systems in the cloud progressively tend to be built as reliable, scalable, and highly available. As more IT businesses adopt cloud-native practices and rely on distributed applications, it becomes essential to have a system that enables continuous monitoring, maintaining, and delivering new systems to production. The vast majority of organizations rely on tools such as Kubernetes \cite{k8s_docs} or HashiCorp Nomad \cite{nomad_docs}, but these often come with high operational costs and require comprehensive technical expertise to maintain them. As an example, to setup a Kubernetes cluster, a company needs to have at least 4 CPU cores and 16 gigabytes of RAM for the control-plane component, and even then, managing upgrades, configuring networking and storage, handling security policies, and troubleshooting failures can be highly complex, often requiring specialized DevOps teams and extensive ongoing maintenance.

The \textit{orchestration principle} \cite{A, ASup} underlies the aforementioned platforms. Orchestration is the process of automated management, coordination, and deployment of multiple services and components in distributed systems to ensure their consistent and efficient operation. Furthermore, due to the nature of the modern orchestration platforms, it becomes problematic to modify them and add new features. As an example, some platforms rely on the push-based model, which could lead to a certain amount of problems like: breaking delivery after network connectivity errors, or workload duplication. For businesses, it could be critical to deal with these issues, since any kind of failure on the infrastructure could potentially lead to high financial losses.

The complexity of orchestration platforms, and potential issues with maintainability and workloads management, create significant barrier for organizations, that cannot afford large DevOps departments or teams. Their setup and maintenance require not only hardware resources but also constant attention from skilled engineers. Furthermore, the reliance on a push-based model limits fault tolerance and creates bottlenecks in distributed scheduling. This thesis addresses related limitations by exploring a pull-based approach \cite{B} to distributed workload management. The key goal of the pull-based approach is not to directly order another service to perform a task, but to place this task in some registry, from which this service will later receive the task and perform it.  So, in practice instead of pushing tasks from a central system, worker nodes independently pull their configuration and workload data from the control plane. This model aims to increase reliability, simplify maintenance, and make distributed orchestration more accessible.

The purpose of this work is designing and implementing a working prototype of a distributed workloads management system with \textit{pull-model approach}. Key objectives are: reducing operational overhead, designing decentralized architecture with ability to scale, improving security by automating manual infrastructure operations and creating a lightweight flexible alternative to existing orchestration platforms. During each phase, artifacts will be collected, so the solution could be replicated.

To achieve the proposed goals, we aim to apply the Design Science Research \cite{D} methodology, which mainly focuses on building and evaluating artifacts to solve real-world problems. This approach emphasizes iterative development, where each stage informs and refines the next, ensuring that the resulting system effectively addresses the identified challenges. The work is separated into several stages:

\begin{enumerate}
\item Problem identification and requirements clarification – analyze existing orchestration solutions, identify their limitations, and define the functional and non-functional requirements for the proposed system.
\item System design and prototype development – develop the architecture and implement a working prototype of the distributed workload management system, incorporating features like task scheduling, fault tolerance, and scalability.
\item Validation and evaluation – test the prototype under controlled and realistic conditions, measure its performance, reliability, and resource efficiency, and compare it against existing platforms to assess improvements and identify areas for further refinement.
\end{enumerate}

This methodology ensures a structured, evidence-based approach to designing and assessing the proposed system while keeping practical applicability as a central focus.

It should be explicitly noted that generative artificial intelligence (GenAI) tools were used in this research solely as a language and editing aid. Specifically, GenAI assisted in refining phrasing, improving grammatical correctness, and enhancing the academic tone of the exposition. It was employed to challenge initial formulations, suggest clearer syntactic structures, and ensure terminological consistency particularly during the drafting and revision of descriptive and expository sections. However, no analytical, conceptual, architectural, or evaluative work was delegated to or generated by AI systems. All problem formulation, requirements engineering, system design decisions, implementation logic, experimental validation, and scholarly interpretation were conducted exclusively by the author. The intellectual contribution, critical reasoning, and technical insights presented in this dissertation are entirely human-derived.

The expected outcome of this research is a lightweight and reliable orchestration system prototype that demonstrates the practical benefits of applying a pull-based model in distributed workload management. The developed system is expected to reduce operational overhead by automating configuration delivery and minimizing manual interventions. It will introduce a decentralized control plane that relies on REST API and PostgreSQL for efficient coordination and data storage. Additionally, the work aims to provide a clear comparison with existing orchestration platforms, highlighting improvements in scalability, fault tolerance, and resource efficiency. As a result, this thesis contributes to simplifying infrastructure management and offers an alternative orchestration approach that can be more easily adopted by organizations with different operational scales.
